---
title: "Exercise 3 with solutions: Regression of categorical and ordinal data"
author: "Stefan Konigorski"
date: "January 6, 2020"
output:
  html_document: default
---

## Load KiGGS data

```{r}
#dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
#load(dat_link)
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
dat <- KiGGS03_06
```

## Logistic, multinomial, ordinal regression

Predict the birth weight of children (e017a.k, in categories) by the BMI of their parents (Mbmi.k, Vbmi.k), their sex (sex), and the birth week (e016z).
For this, use a multinomial regression and ordinal regression model, and compare the results to pairwise logistic regression models. 
Are the results similar? 
What are the advantages/disadvantages of each approach?

```{r}
# Preparation - check variables (eg if they are factor variables, and if yes, how many categories they have), especially important for e017a.k
str(dat$e017a.k) # -> factor with 6 categories
str(dat$Mbmi.k)
str(dat$Vbmi.k)
str(dat$sex)
str(dat$e016z)

# Format variables
dat$e017a.k <- factor(dat$e017a.k, labels = c("<2500", "2500-2999", "3000-3499", "3500-3999", "4000-4499", ">4500"))
dat$Mbmi.k  <- factor(dat$Mbmi.k, labels = c("underweight", "normal weight", "overweight", "obese"))
dat$Vbmi.k  <- factor(dat$Vbmi.k, labels = c("underweight", "normal weight", "overweight", "obese"))
dat$sex     <- factor(dat$sex, labels = c("boys", "girls"))
dat$e016z   <- as.numeric(as.character(dat$e016z))

# Look at variables
table(dat$e017a.k) 
table(dat$Mbmi.k)
table(dat$Vbmi.k)
table(dat$sex)
table(dat$e016z)
```

First, compute the models only using e016z as predictor

```{r}
# Multinomial regression:
# install.package("nnet")
library(nnet)
levels(dat$e017a.k)
fit_mr1 <- nnet::multinom(e017a.k ~ e016z, data = dat)

# Look at results: 
summary(fit_mr1) # -> one coefficient (for each predictor) for the respective comparison of the category (eg 2500-2999) with the reference category (here: <2500)

# p-values for hypothesis tests of the regression coefficients (whether they are 0 or not) have to be computed by hand, e.g.:
W <- summary(fit_mr1)$coefficients/summary(fit_mr1)$standard.errors
(1 - pnorm(abs(W), 0, 1)) * 2

# Compute odds ratio:
exp(summary(fit_mr1)$coefficients)

# Predicted probabilities: 
head(fitted(fit_mr1))
```

```{r}
# Ordinal regression:
# install.package("MASS")
library(MASS)
fit_or1 <- MASS::polr(e017a.k ~ e016z, method = "logistic", Hess = TRUE, data = dat)

# Look at results: 
summary(fit_or1) # -> Look under 'Coefficients': one coefficient (for each predictor) which contains the effect of the birth week on the birth weight category (under the proportional odds assumption)

# p-values:
pnorm(abs(summary(fit_or1)$coefficients[, 3]), lower.tail = FALSE) * 2

# Compute odds ratio:
exp(summary(fit_or1)$coefficients[, 1])
```

```{r}
# Pairwise logistic regression:
dat_kat12 <- dat[dat$e017a.k %in% c("<2500", "2500-2999"), ]
dat_kat13 <- dat[dat$e017a.k %in% c("<2500", "3000-3499"), ]
dat_kat14 <- dat[dat$e017a.k %in% c("<2500", "3500-3999"), ]
dat_kat15 <- dat[dat$e017a.k %in% c("<2500", "4000-4499"), ]
dat_kat16 <- dat[dat$e017a.k %in% c("<2500", ">4500"), ]
fit_lr12 <- glm(e017a.k ~ e016z, family = binomial(link = "logit"), data = dat_kat12)
fit_lr13 <- glm(e017a.k ~ e016z, family = binomial(link = "logit"), data = dat_kat13)
fit_lr14 <- glm(e017a.k ~ e016z, family = binomial(link = "logit"), data = dat_kat14)
fit_lr15 <- glm(e017a.k ~ e016z, family = binomial(link = "logit"), data = dat_kat15)
fit_lr16 <- glm(e017a.k ~ e016z, family = binomial(link = "logit"), data = dat_kat16)

# look at results:
summary(fit_lr12)$coefficients
summary(fit_lr13)$coefficients
summary(fit_lr14)$coefficients
summary(fit_lr15)$coefficients
summary(fit_lr16)$coefficients

# odds ratio with 95% CI:
library(jtools)
jtools::summ(fit_lr12, exp = T, confint = T, model.fit = F, digits = 3)
```

Now incorporate more predictors:

```{r}
# Multinomial regression:
fit_mr2 <- nnet::multinom(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), data = dat)
summary(fit_mr2)
```

```{r}
# Ordinal regression:
fit_or2 <- MASS::polr(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), method = "logistic", Hess = TRUE, data = dat)
summary(fit_or2) 
```

```{r}
# Pairwise logistic regression:
fit_lr12 <- glm(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), family = binomial(link = "logit"), data = dat_kat12)
fit_lr13 <- glm(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), family = binomial(link = "logit"), data = dat_kat13)
fit_lr14 <- glm(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), family = binomial(link = "logit"), data = dat_kat14)
fit_lr15 <- glm(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), family = binomial(link = "logit"), data = dat_kat15)
fit_lr16 <- glm(e017a.k ~ Mbmi.k + Vbmi.k + sex + as.numeric(e016z), family = binomial(link = "logit"), data = dat_kat16)
summary(fit_lr12)$coefficients
summary(fit_lr13)$coefficients
summary(fit_lr14)$coefficients
summary(fit_lr15)$coefficients
summary(fit_lr16)$coefficients
```
