---
title: "Exercise 2 with solutions: Regression with count data"
author: "Stefan Konigorski"
date: "January 6, 2020"
output:
  html_document: default
---

## Load KiGGS data

```{r}
#dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
#load(dat_link)
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
dat <- KiGGS03_06
```

## Poisson und Negative Binomial Regression

Predict the amount of visits to the children's doctor (E085z01) by the number of siblings (e006B1), the sex and age of the children (sex, age2), place of residence (STALA, OW) and the monthly household income (e093), using a Poisson regression model. Interpret the results. Which variables are associated with the outcome? Is the model a good fit to the data?

```{r}
# Preparation - check variables (eg if they are factor variables, and if yes, how many categories they have), especially important for E085z01
str(dat$E085z01)
str(dat$e006B1)
str(dat$sex)
str(dat$age2)
str(dat$STALA)
str(dat$OW)
str(dat$e093)
```

```{r}
# Format variables
dat$E085z01 <- as.numeric(as.character(dat$E085z01))
dat$e006B1<- factor(dat$e006B1, labels = c("no sibling", "1 sibling", "2 or more siblings"))
dat$sex   <- factor(dat$sex, labels = c("boys", "girls"))
dat$age2  <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
dat$OW    <- factor(dat$OW, labels = c("east", "west"))
dat$STALA <- factor(dat$STALA, labels = c("rural", "small_city", "medium_city", "urban"))
```

```{r}
# Look at variables
table(dat$E085z01)
table(dat$e006B1)
table(dat$sex)
table(dat$age2)
table(dat$STALA)
table(dat$OW)
table(dat$e093)
```

```{r}
# look at distribution of outcome
hist(dat$E085z01)
mean(dat$E085z01, na.rm = TRUE)
var(dat$E085z01, na.rm = TRUE)
# -> mean and var are very different (var >> mean since long tail on the right), i.e. we should rather compute a Negative Binomial regression model! Still, we will fit a Poisson regression model here for illustration sake 
```

```{r}
# Poisson Regression: first predict dat$E085z01 only by sex:
fit1 <- glm(dat$E085z01 ~ sex, family = poisson(link = "log"), data = dat)

# results:
summary(fit1)

# for interpretation: e^ to have interpretation on the scale of frequencies (and not of log-frequencies):
exp(coef(fit1))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit1, exp = T, confint = T, model.fit = F, digits = 3)
```

```{r}
# Predicted frequencies by the model:
pred_1 <- predict(fit1, type="response")

# Compare these with the observed frequencies:
predictions_1 <- data.frame(obs = dat$E085z01[1:17640 %in% names(pred_1)], pred_1 = pred_1)
cor(predictions_1$obs, predictions_1$pred_1)
plot(predictions_1$obs, predictions_1$pred_1)

# -> the model doesn't seem to predict anything - sex is only very weakly associated with the amount of visits of the children's doctor, i.e. the model predicts about the same number of doctor visits for every child (mean number of visits for boys/girls: 2.87 and 2.81).
```

```{r}
# fit model with all variables:
fit2 <- glm(dat$E085z01 ~ e006B1 + sex + age2 + STALA + OW + e093, family = poisson(link = "log"), data = dat)

# results:
summary(fit2)
exp(coef(fit2))
library(jtools)
jtools::summ(fit2, exp = T, confint = T, model.fit = F, digits = 3)
```

```{r}
# Predictions:
predictions_2 <- data.frame(emp = dat$E085z01[1:17640 %in% names(predict(fit2, type="response"))], exp = predict(fit2, type="response"))
cor(predictions_2$emp, predictions_2$exp)
plot(predictions_2$emp, predictions_2$exp)
# -> better, but (at least graphically) still not really convincing.
```
