---
title: "Exercise 3: Regression of categorical and ordinal data"
author: "Stefan Konigorski"
date: "January 6, 2020"
output:
  html_document: default
---

## Load KiGGS data

```{r}
#dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
#load(dat_link)
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
dat <- KiGGS03_06
```

## Logistic, multinomial, ordinal regression

Predict the birth weight of children (e017a.k, in categories) by the BMI of their parents (Mbmi.k, Vbmi.k), their sex (sex), and the birth week (e016z).
For this, use a multinomial regression and ordinal regression model, and compare the results to pairwise logistic regression models. 
Are the results similar? 
What are the advantages/disadvantages of each approach?

```{r}
# Preparation - check variables (eg if they are factor variables, and if yes, how many categories they have), especially important for e017a.k
str(dat$e017a.k)
str(dat$Mbmi.k)
str(dat$Vbmi.k)
str(dat$sex)
str(dat$e016z)

# Format variables
dat$e017a.k <- factor(dat$e017a.k, labels = c("<2500", "2500-2999", "3000-3499", "3500-3999", "4000-4499", ">4500"))
dat$Mbmi.k  <- factor(dat$Mbmi.k, labels = c("underweight", "normal weight", "overweight", "obese"))
dat$Vbmi.k  <- factor(dat$Vbmi.k, labels = c("underweight", "normal weight", "overweight", "obese"))
dat$sex     <- factor(dat$sex, labels = c("boys", "girls"))
dat$e016z   <- as.numeric(as.character(dat$e016z))

# Look at variables
table(dat$e017a.k)
table(dat$Mbmi.k)
table(dat$Vbmi.k)
table(dat$sex)
table(dat$e016z)
```

```{r}
# Multinomial regression:
library(nnet)

# Ordinal regression:
library(MASS)

# Pairwise logistic regression:
...
```


