---
title: "Exercise 1: Logistic regression"
author: "Stefan Konigorski"
date: "January 6, 2020"
output:
  html_document: default
---

## Load KiGGS data

```{r}
#dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
#load(dat_link)
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
dat <- KiGGS03_06
```

## Logistic regression

a. Investigate if the amount and frequency of eating pancakes (fq41 and fq41a) are associated with shortsightedness (e0251), using logistic regression. (Question: What is predicted: incidence or prevalence of shortsightedness?). Interpret the results.

b. Add further predictors and interpret the results. Also look at the predicted probabilities by the model and the classification table of true vs. predicted shortsightedness.

```{r}
# Preparation for logistic regression - check variables (eg if they are factor variables, and if yes, how many categories they have), especially e0251
str(dat$e0251)
str(dat$fq41)
str(dat$fq41a)
table(dat$e0251)
table(dat$fq41)
table(dat$fq41a)

# format variables
dat$fq41a <- factor(dat$fq41a, labels = c("never", "1/4 piece or less", "1/2 piece", "1 piece", "2 pieces", "3 pieces or more"))
dat$fq41a <- droplevels(dat$fq41a)
dat$fq41 <- factor(dat$fq41, labels = c("never", "1x/month", "2-3x/month", "1-2x/week", "3-4x/week", "5-6x/week", "1x/day", "2-3x/day", "4-5x/day"))
dat$e0251 <- factor(dat$e0251, labels = c("yes", "no", "don't know"))

# Compute a logistic regression with these variables and interpret the results.

# Logistic regression in R needs Y with values 0, 1 or a factor with two levels, and predicts the probability of "1", or the probability of the factor with higher level.
# Transform the outcome variable for a better interpretation of the results. Remove the values "don't know" and the factor level to create a binary variable. Then reorder the levels.
# Compute a logistic regression with this transformed outcome variable.
```
