---
title: "Homework 12"
author: "Juan Carlos Niño"
date: "January 26, 2020"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Linear mixed model for hierarchical data

Load the HRS dataset and do the following preparations:

```{r}
library(readr)
dat_link <- url("https://www.dropbox.com/s/yox8fvt91ff73ye/hrs.RData?dl=1")
load(dat_link)
#load("C:/Users/stefan.konigorski/Desktop/data_Teaching/HRS/CogImp/hrs.RData")

# Preparations
# rename variable at first timepoint
names(hrs)[5] <- "R1IMRC"
# define variable at second timepoint - Integrate the variables R2AIMRC10 and R2HIMRC20, which both capture different subgroups of the sample at the second timepoint.
table(is.na(hrs$R2AIMRC10)) # -> only ~7000 of ~36000 observations
table(is.na(hrs$R2HIMRC20)) # -> only ~10000 of ~36000 observations
table(!is.na(hrs$R2HIMRC20) & !is.na(hrs$R2AIMRC10)) # -> no person has a value at both variables, i.e. the variables complement each other
hrs$R2IMRC <- rowSums( cbind (hrs$R2AIMRC10, hrs$R2HIMRC20), na.rm=TRUE)
hrs$R2IMRC[!(!is.na(hrs$R2HIMRC20) | !is.na(hrs$R2AIMRC10))] <- NA
table(is.na(hrs$R2IMRC))
# Rescale R1IMRC and R2IMRC onto the same scale as all other variables (measures 3-12 were on a scale of 1-10, measures 1-2 on the scale 1-20)
hrs$R1IMRC <- hrs$R1IMRC/2
hrs$R2IMRC <- hrs$R2IMRC/2

# Extract variables for first investigation
data_IMRC <- data.frame(HHID = hrs$HHID, PN = hrs$PN, HHID_PN = paste(hrs$HHID, hrs$PN, sep = "_"),
                        T1 = hrs$R1IMRC, T2 = hrs$R2IMRC, T3 = hrs$R3IMRC, T4 = hrs$R4IMRC, T5 = hrs$R5IMRC, T6 = hrs$R6IMRC,
                        T7 = hrs$R7IMRC, T8 = hrs$R8IMRC, T9 = hrs$R9IMRC, T10 = hrs$R10IMRC, T11 = hrs$R11IMRC, T12 = hrs$R12IMRC)
```

Answer the following questions: 

a) How many observations are there in the dataset?

```{r}

dim(hrs)
#there are 36167 observations

```


b) How many households are there in the dataset (observations are clustered in households)?

```{r}

str(hrs$HHID)
table(table(hrs$HHID))
10570+11775+568+73+9+1
#there are 22996 households

```



c) Now, compute a linear regression model to predict T2 by T1, i.e. the cognitive performance at the second timepoint by the cognitive performance at the first timepoint. Also, compute two linear mixed models, one with random intercept, one with random intercept+slope to predict T2 by T1. Compare the results, i.e. the estimated effect of T1 on T2. Are they similar, or are they very different?

```{r}

#linear regression model
fit1 <- lm (T2 ~ T1, data = data_IMRC)
summary(fit1)

# Prediction:
pred.fit1 <- predict(fit1)
head(pred.fit1)

```




```{r}
#linear mixed model
library(nlme)
fit2 <- lme (T2 ~ T1, random= ~ 1|HHID, data = data_IMRC, na.action = na.exclude)
summary(fit2)

# Prediction:
pred.fit2 <- predict(fit2)
head(pred.fit2)

```

```{r}
#linear mixed model
library(nlme)
fit3 <- lme (T2 ~ T1, random= ~ T1|HHID, data = data_IMRC, na.action = na.exclude)
summary(fit3)

# Prediction:
pred.fit3 <- predict(fit3)
head(pred.fit3)

```

The 3 different models maintain the more or less the prediciton of time with 0.53 as the first 2 significance digits, then changing the third number over the models.



## Exercise 2: Linear mixed model for longitudinal data

Go through R_12b_example_3.Rmd, run the R code, and look at the results. Write down 2 things that you learned by doing this, or 2 things that you found interesting. Also write down 2 questions on things that have not been clear to you in the modeling, or that you want to further investigate.


#Learned
-Do the plotting of this dataset is hard without transforming the format of the data
-Despite being alwasy the same data the clustering alters dramatically the results

#Questions
-the correlation part of the function models ......HHID_PN, correlation = corAR1( form = ~ 1 | HHID_PN), na.a......
-why not the use of context variables?









