---
title: "Linear mixed models in the MathAchieve dataset"
author: "Stefan Konigorski"
date: "January 20, 2020"
output:
  html_document: default
---

Note: this is a continuation of the exercise in R_11b_exercise_LMM.Rmd

# Load and inspect MathAchieve dataset

```{r}
# install.packages("nlme")
library(nlme)
```

## Model with SES, Minority, Sex as fixed-effects predictors

```{r}
fit3 <- lm(MathAch ~ SES + Sex + Minority, data = MathAchieve)
summary(fit3)

pred3 <- predict(fit3)
plot(pred3, MathAchieve$MathAch, xlab = "Predicted MathAch values", ylab = "Actual MathAch values")
```

## Does school have an effect?

```{r}
head(MathAchieve)
str(MathAchieve)

length(levels(MathAchieve$School))

# Look at association of SES and MathAch in different schools
MathAchieve1 <- MathAchieve[MathAchieve$School %in% levels(MathAchieve$School)[1:16], ]
MathAchieve2 <- MathAchieve[MathAchieve$School %in% levels(MathAchieve$School)[17:32], ]

library(ggplot2)
ggplot(MathAchieve1, aes(x = SES, y = MathAch)) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap( . ~ School, drop = FALSE)

ggplot(MathAchieve2, aes(x = SES, y = MathAch)) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap( . ~ School, drop = FALSE)
```

How can we consider school? -> Use mixed models

## Mixed models: random intercept

```{r}
library(nlme)
fit5a <- lme(MathAch ~ SES + Sex + Minority, random = ~1|School, data = MathAchieve)
summary(fit5a)

library(lme4)
fit5b <- lmer(MathAch ~ SES + Sex + Minority + (1|School), data = MathAchieve)
summary(fit5b)

# Compare estimates of the fixed effects:
summary(fit5a)$tTable
summary(fit5b)$coef
```

## Mixed models: random intercept, random slope

```{r}
fit6a <- lme(MathAch ~ SES + Sex + Minority, random = ~SES|School, data = MathAchieve)
summary(fit6a)

fit6b <- lmer(MathAch ~ SES + Sex + Minority + (SES|School), data = MathAchieve)
summary(fit6b)
```

## Mixed models: ICC

```{r}
#install.packages("psychometric")
library(psychometric)
ICC1.lme(dv = MathAch, grp = School, data = MathAchieve)

# by hand:
summary(lme(MathAch ~ 1, random = ~1|School, data = MathAchieve))
2.934966^2 / (2.934966^2 + 6.256862^2)
```

## Incorporating context variables (= macro variables)

```{r}
fit7 <- lme(MathAch ~ SES + Sex + Minority + MEANSES, random = ~1|School, data = MathAchieve)
summary(fit7)
```

## Predict school specific effects of SES on MathAch

```{r}
library(ggplot2)

###################################################################
# Model 1:

fit8a <- lme(MathAch ~ SES, random = ~1|School, data = MathAchieve)
pred.fit8a <- predict(fit8a)

summary(fit8a)
head(MathAchieve)
head(pred.fit8a)

# Prediction of first student just using fixed effects:
12.657480 + 2.390196 * (-1.528)
# the difference between this and the predicted value from predic() is the random intercept term for the first school
```

```{r}
# Different visualizations:

ggplot(MathAchieve, aes(x = SES, y = MathAch)) +
  geom_point(size = 0.1) +
  geom_line(aes(y = pred.fit8a, group = School))

ggplot(MathAchieve, aes(x = SES, y = MathAch)) +
  geom_line(aes(y = pred.fit8a, group = School, colour = School)) + 
  theme(legend.position = "none")
```


```{r}
###################################################################
# Model 2:

fit8b <- lme(MathAch ~ SES, random = ~SES|School, data = MathAchieve)
pred.fit8b <- predict(fit8b)

ggplot(MathAchieve, aes(x = SES, y = MathAch)) +
  geom_line(aes(y = pred.fit8b, group = School, colour = School)) + 
  theme(legend.position = "none")

# -> compare with the plots of the actual math grades of the single schools!
# this shows the effect of the weighting of the overall effect and the school-specific observations
```

```{r}
###################################################################
# Model 3:
fit8c <- lme(MathAch ~ SES + MEANSES, random = ~SES|School, data = MathAchieve)
pred.fit8c <- predict(fit8c)

ggplot(MathAchieve, aes(x = SES, y = MathAch, colour = MEANSES)) +
  geom_line(aes(y = pred.fit8c, group = School)) 
```
