---
title: "Class 2 - exercises"
author: "Stefan Konigorski"
date: "October 28, 2019"
output:
  html_document: default
---

## Exercise 1: Practise working with R Markdown

Inspect the R Markdown file R_1_homework_solution.Rmd and try to edit and knit it yourself to generate a html or Word output file.

## Exercise 2: Check Pima diabetes dataset

a) Run the following commands and inspect the results, in terms of what the functions are doing, and if the data values make sense and seem correct.

```{r}
# Load/import data (edit so that the path points to the file and folder where you have saved the Pima_diabetes.Rdata file on your laptop)
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/PimaIndians/Pima_diabetes.RData")

# Alternatively: import the file from the following dropbox link:
# Pima_diabetes <- read.csv(file = url("https://www.dropbox.com/s/4s4rhf6abda6gw2/Pima_diabetes.csv?dl=1"))

# Check data frame in general
str(Pima_diabetes)
summary(Pima_diabetes)
dim(Pima_diabetes)
# how many observations:
nrow(Pima_diabetes)
# how many variables:
ncol(Pima_diabetes) 

# check single variables
table(Pima_diabetes$Glucose)
table(Pima_diabetes$Glucose == 0)
table(Pima_diabetes$BMI < 10)
table(Pima_diabetes$BMI <= 10)

# more complicated checks
table(!Pima_diabetes$Glucose == 0)
table(Pima_diabetes$Glucose == 0 | Pima_diabetes$BMI == 10)
table(Pima_diabetes$Glucose == 0 & Pima_diabetes$BMI == 10)
```

b) Answer the following questions (these are already computing some descriptive statistics)

```{r}
# How many women in the sample have diabetes?

# How many women are older than 50 years?

# Do all insulin values make sense?
```

c) Do some transformations (do at home)

```{r}
# check Glucose levels
table(!Pima_diabetes$Glucose == 0)

# transform variables
Pima_diabetes$Glucose[Pima_diabetes$Glucose == 0] <- NA

# check again:
table(!Pima_diabetes$Glucose == 0)
Pima_diabetes$Glucose[!is.na(Pima_diabetes$Glucose)]
```

d) Some small tidyverse example (do at home)

```{r}
# install.packages(dplyr)
library(dplyr)

# Compute average BMI of those women with more than 10 pregnancies
Pima_diabetes %>%
	dplyr::select(Pregnancies, BMI) %>% 
	dplyr::filter(Pregnancies > 10) %>%
	dplyr::summarize(avg_BMI_high_Preg = mean(BMI), n = n())

# Compute average BMI of those women with at most 5 pregnancies
Pima_diabetes %>%
	dplyr::select(Pregnancies, BMI) %>% 
	dplyr::filter(Pregnancies <= 5) %>%
	dplyr::summarize(avg_BMI_few_Preg = mean(BMI), n = n())
```

## Exercise 3: Transform NoShow dataset

a) Start with some exploration of the dataset

```{r}
library(readr)
dat <- readr::read_csv(file = "C:/Users/stefan.konigorski/Desktop/data_Teaching/NoShow/KaggleV2-May-2016.csv")

# How many levels of gender are there?

# How many hypertension cases are there?

# do all age values make sense?
```

Run the following commands and inspect the results. 

```{r}
# Import data
dat <- read.csv(file = "C:/Users/stefan.konigorski/Desktop/data_Teaching/NoShow/KaggleV2-May-2016.csv")

# transform variables
dat$Age[dat$Age < 0] <- NA
# check again:
table(dat$Age < 0)

# another way how to check how many observations there are in the dataset
length(dat$PatientId)
# do all patients only come once?
length(unique(dat$PatientId))

# transform dataset (extract subset)
dat_complete1 <- na.omit(dat)
dat_complete2 <- dat[complete.cases(dat), ]
dat_complete3 <- dat[!is.na(dat$Age), ]

# more examples for transformation:
d <- data.frame(x = c(1:4), y = c(NA, NA, 2, 4))
d
na.omit(d)
d[complete.cases(d),]
```

## Exercise 4: times and dates

a) Practise to work with dates and times in R.

```{r}
# Download NoShow csv file KaggleV2-May-2016.csv from Blackboard and import in R using both the read.csv and readr::read_csv function.
library(readr)
dat1 <- readr::read_csv(file = "C:/Users/stefan.konigorski/Desktop/data_Teaching/NoShow/KaggleV2-May-2016.csv")
dat2 <- read.csv(file = "C:/Users/stefan.konigorski/Desktop/data_Teaching/NoShow/KaggleV2-May-2016.csv")

# look at data frames
head(dat1)
head(dat2)

str(dat1)
str(dat2)
# -> readr::read_csv has already read in the date/time variables ScheduledDay, AppointmentDay correctly as date/time variable
# -> read.csv has read in the date/time variables ScheduledDay, AppointmentDay as factors

# Use as.Date() function to create variables of class Date:
SDay_1 <- as.Date(dat1$ScheduledDay)
ADay_1 <- as.Date(dat1$AppointmentDay)
SDay_2 <- as.Date(dat2$ScheduledDay)
ADay_2 <- as.Date(dat2$AppointmentDay)

# check classes
class(dat1$ScheduledDay)
class(dat2$ScheduledDay)
class(SDay_1)
class(SDay_2)

# compare:
head(dat1$ScheduledDay)
head(SDay_1)
head(SDay_2)

head(dat1$AppointmentDay)
head(ADay_1)
head(ADay_2)

# now look at their values as numbers:
head(as.numeric(dat1$ScheduledDay)) # -> seconds since 1970-1-1
head(as.numeric(SDay_1)) # -> days since 1970-1-1
head(as.numeric(dat2$ScheduledDay)) # not a number that makes sense for calculations (rank in ordering of dates)

# Now you can e.g. make calculations with these variables, e.g. create new variable that is the difference between two dates
dat1$hour_diff <- dat1$ScheduledDay - dat1$AppointmentDay
str(dat1$hour_diff)

# Now answer the question if the AppointmentDay is always before the ScheduledDay?
```

b) Check the variables ScheduledDay and AppointmentDay. Is the AppointmentDay always before the ScheduledDay? If something seems weird, do further checks.

Variable description from https://www.kaggle.com/joniarroba/noshowappointments: 

* DataMarcacaoConsulta (from the order in the dataset: ScheduledDay) = The day of the actual appointment, when they have to visit the doctor.
* DataAgendamento (from the order in the dataset: AppointmentDay) = The day someone called or registered the appointment, this is before appointment of course.

```{r}
```


