---
title: "Class 2 - solutions to exercises"
author: "Stefan Konigorski"
date: "October 28, 2019"
output:
  html_document: default
---

## Exercise 2: Check data set

b) Answer the following questions (these are already computing some descriptive statistics)

```{r}
# Load/import data
load("C:/Users/stefan.konigorski/Desktop/data_Teaching/PimaIndians/Pima_diabetes.RData")
# Pima_diabetes <- read.csv(file = url("https://www.dropbox.com/s/4s4rhf6abda6gw2/Pima_diabetes.csv?dl=1"))

# How many women in the sample have diabetes? (outcome==1)
table(Pima_diabetes$Outcome)

# How many women are older than 50 years?
table(Pima_diabetes$Age>50)

# Do all insulin values make sense?
summary(Pima_diabetes$Insulin)
```

## Exercise 3: times and dates

a) Start with some exploration of the dataset

```{r}
library(readr)
dat <- readr::read_csv(file = "C:/Users/stefan.konigorski/Desktop/data_Teaching/NoShow/KaggleV2-May-2016.csv")

# How many levels of gender are there?
table(dat$Gender)

# How many hypertension cases are there? (21801)
table(dat$Hipertension)

# do all age values make sense?
table(dat$Age)
hist(dat$Age)
table(dat$Age)
table(dat$Age == 0)
table(dat$Age < 0)
summary(dat$Age)
# -> 1x value -1 which cannot be correct. 3539x value 0, which might be correct.
```

## Exercise 4: times and dates

b) Check the variables ScheduledDay and AppointmentDay. Is the AppointmentDay always before the ScheduledDay? If something seems weird, do further checks.

Variable description from https://www.kaggle.com/joniarroba/noshowappointments: 

* DataMarcacaoConsulta (from the order in the dataset: ScheduledDay) = The day of the actual appointment, when they have to visit the doctor.
* DataAgendamento (from the order in the dataset: AppointmentDay) = The day someone called or registered the appointment, this is before appointment of course.

```{r}
# Question: is the AppointmentDay always before the ScheduledDay
table(dat$ScheduledDay > dat$AppointmentDay)
table(dat$ScheduledDay < dat$AppointmentDay)
table(dat$ScheduledDay == dat$AppointmentDay)

dat$hour_diff <- dat$ScheduledDay - dat$AppointmentDay
table(dat$hour_diff > 0)

# and look at it:
hist(as.numeric(dat$hour_diff), main = "Histogram of difference between \n scheduling and appointment day in hours", xlab = "Difference in hours")
hist(as.numeric(dat$hour_diff)/24, main = "Histogram of difference between \n scheduling and appointment day in days", xlab = "Difference in days")

# some further checks: what is going on?
head(dat$ScheduledDay)
head(dat$AppointmentDay)
# -> ah, maybe the two are wrongly labeled (translated?!), that ScheduledDay is when the appointment was made, and AppointmentDay is when the patient saw the doctor
table(dat$AppointmentDay > dat$ScheduledDay)

# next thought: AppointmentDay was always set to 0:00 o'clock. Set it to the next day i.e. add 1 day, then check again:
dat$AppointmentDay_cor <- dat$AppointmentDay + 86400
table(dat$AppointmentDay_cor > dat$ScheduledDay)

# -> so this seems to be the right thing: (i) variables were mixed up, (ii) 1 day should be added, (iii) then three values corrected:
dat[!(dat$AppointmentDay_cor > dat$ScheduledDay), ]
# what to do with these?
# depending on the analysis, set maybe to NA since don't know what is wrong
dat$AppointmentDay_cor[!(dat$AppointmentDay_cor > dat$ScheduledDay)] <- NA
dat$ScheduledDay[!(dat$AppointmentDay_cor > dat$ScheduledDay)] <- NA

#check again
table(dat$AppointmentDay_cor > dat$ScheduledDay)
# dat[!(dat$AppointmentDay_cor > dat$ScheduledDay), ]
```
