---
title: "Homework 3 - Solutions"
author: "Stefan Konigorski"
date: "November 4, 2019"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Compute frequencies in the Pima diabetes dataset

Load the Pima diabetes dataset:

```{r}
load(file = url("https://www.dropbox.com/s/j50z4e17l5aabk8/Pima_diabetes.RData?dl=1"))
```

**Task:** Which variables are measured on a nominal level?

**Solution:**

```{r}
head(Pima_diabetes)
str(Pima_diabetes)
```

Only the outcome is nominal!

**Task:** Now compute frequency tables, barplots, and mosaic plots of all nominal variables in the dataset.

**Solution:**

```{r}
# to get better interpretable plots and tables with labels of the variable, transform the variable to a 'factor' and add labels
# annotation: 0 - does not have diabetes; 1 - has diabetes
Pima_diabetes$Diabetes <- factor(Pima_diabetes$Outcome, labels = c("no", "yes"))

# Table of absolute frequencies:
table(Pima_diabetes$Outcome)
table(Pima_diabetes$Diabetes)
table_diab <- table(Pima_diabetes$Diabetes)

# Table with relative frequencies:
table(Pima_diabetes$Diabetes) / nrow(Pima_diabetes)

# barplot
barplot(table_diab, xlab = "Diabetes", ylab = "Absolute frequencies")

# mosaic plot
mosaicplot(table_diab, main = "Mosaic plot of Diabetes", color = TRUE)
# -> the frequencies are much less clearly represented here compared to a barplot.
```

**Task:** Next, create a variable which describes whether a woman had more or less than 4 pregnancies. Then, use this variable to create a 2x2 table with diabetes outcome. Do you see an indication of whether the number of pregnancies is associated with diabetes prevalence? Do you think your investigation is a good way to investigate this?

**Solution:**
```{r}
Pima_diabetes$Preg4 <- 0
Pima_diabetes$Preg4[Pima_diabetes$Pregnancies >= 4] <- 1 # note you have to choose and assign women with 4 pregnancies to either group
Pima_diabetes$Preg4 <- factor(Pima_diabetes$Preg4, labels = c("< 4", ">= 4"))
table(Pima_diabetes$Preg4, Pima_diabetes$Diabetes, dnn = c("Number of pregnancies", "Diabetes"))

```

## Exercise 2: Generate a table with descriptive statistics

Use any dataset (a dataset that you have worked with in the past, or that you are currently working with, a dataset that is available on Blackboard, in R or that you have downloaded from the internet), and generate a table with descriptive statistics of the main variables of interest.

## Exercise 3: Plots using ggplot2

Load the NoShow dataset:

```{r}
load(file = url("https://www.dropbox.com/s/32pyh375kjkqv14/NoShowdata.RData?dl=1"))
head(NoShowdata)
```

Use ggplot2 to generate the following plots:

- Create a boxplots of Age (stratified) by neighborhood.
- Create a histogram of Age.
- Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.
- Stratify this plot further by gender.

What insights can you get from these plots? For which goal would you create these plots?

```{r}
# Prep: load ggplot2 package and format variables
library(ggplot2)
# correct the one negative value of age, and set it to NA
NoShowdata$Age[NoShowdata$Age < 0] <- NA
NoShowdata$Gender <- factor(NoShowdata$Gender, labels = c("female", "male"))
NoShowdata$Diabetes <- factor(NoShowdata$Diabetes, labels = c("FALSE", "TRUE"))
NoShowdata$Noshow <- factor(NoShowdata$"No-show", labels = c("did show up", "did not show up"))
NoShowdata$Neighbourhood <- factor(NoShowdata$Neighbourhood)

# Boxplots of age by neighborhood
ggplot(data = NoShowdata, aes(y = Age)) + geom_boxplot()
ggplot(data = NoShowdata, aes(y = Age, x = Neighbourhood)) + geom_boxplot()

# Histogram of age
ggplot(data = NoShowdata, mapping = aes(x = Age)) + geom_histogram()

# Histogram of age stratified by NoShow
ggplot(data = NoShowdata, mapping = aes(x = Age, col = Noshow)) + geom_histogram()
ggplot(data = NoShowdata, mapping = aes(x = Age, fill = Noshow)) + geom_histogram()
ggplot(data = NoShowdata, mapping = aes(x = Age)) + geom_histogram() + facet_grid(~ Noshow)

# Histogram of age stratified by NoShow and Gender
ggplot(data = NoShowdata, mapping = aes(x = Age, fill = Noshow)) + geom_histogram() + facet_grid(~ Gender)
```
