---
title: "Homework 3"
author: "Stefan Konigorski"
date: "November 4, 2019"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Compute frequencies in the Pima diabetes dataset

Load the Pima diabetes dataset:

```{r}
load(file = url("https://www.dropbox.com/s/j50z4e17l5aabk8/Pima_diabetes.RData?dl=1"))
```

Which variables are measured on a nominal level?

Now compute frequency tables, barplots, and mosaic plots of all nominal variables in the dataset.

Next, create a variable which describes whether a woman had more or less than 4 pregnancies. Then, use this variable to create a 2x2 table with diabetes outcome. Do you see an indication of whether the number of pregnancies is associated with diabetes prevalence? Do you think your investigation is a good way to investigate this?

## Exercise 2: Generate a table with descriptive statistics

Use any dataset (a dataset that you have worked with in the past, or that you are currently working with, a dataset that is available on Blackboard, in R or that you have downloaded from the internet), and generate a table with descriptive statistics of the main variables of interest.

## Exercise 3: Plots using ggplot2

Load the NoShow dataset:

```{r}
load(file = url("https://www.dropbox.com/s/32pyh375kjkqv14/NoShowdata.RData?dl=1"))
head(NoShowdata)
```

Use ggplot2 to generate the following plots:

- Create a boxplots of Age (stratified) by neighborhood.
- Create a histogram of Age.
- Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.
- Stratify this plot further by gender.

What insights can you get from these plots? For which goal would you create these plots?