---
title: "Exercises 3b"
author: "Stefan Konigorski"
date: "November 4, 2019"
output:
  pdf_document: default
---

# Exercise 1: Tables

```{r}
# Load data
load(file = url("https://www.dropbox.com/s/j50z4e17l5aabk8/Pima_diabetes.RData?dl=1"))
Pima_diabetes$Diabetes <- factor(Pima_diabetes$Outcome, labels = c("no", "yes"))
```

Now create custom tables using the qwraps2 package.

Note: need the **results = "asis"** statement when starting the R chunk so that the tables that are specially formatted by the **qwraps2** package are displayed correctly in the knitted output file. 

First small example table:

```{r, results = "asis"}
# install.packages("qwraps2")
library(qwraps2)
options(qwraps2_markup = "markdown")

our_summary0 <-
  list("Diabetes" =
       list("yes" = ~ qwraps2::n_perc0(.data$Diabetes == "yes", show_symbol = TRUE, na_rm = TRUE),
            "no"  = ~ qwraps2::n_perc0(.data$Diabetes == "no", show_symbol = TRUE, na_rm = TRUE))
       )
summary_table(Pima_diabetes, our_summary0)
```

Now a longer table with more details: 

```{r, results = "asis"}
our_summary1 <-
  list("Diabetes" =
       list("yes" = ~ qwraps2::n_perc0(.data$Diabetes == "yes", show_symbol = TRUE, na_rm = TRUE),
            "no"  = ~ qwraps2::n_perc0(.data$Diabetes == "no", show_symbol = TRUE, na_rm = TRUE)),
       "Number of pregnancies" =
       list("0" = ~ qwraps2::n_perc0(.data$Pregnancies == 0, show_symbol = TRUE, na_rm = TRUE),
            "1" = ~ qwraps2::n_perc0(.data$Pregnancies == 1, show_symbol = TRUE, na_rm = TRUE),
            "2"  = ~ qwraps2::n_perc0(.data$Pregnancies == 2, show_symbol = TRUE, na_rm = TRUE),
            "3"  = ~ qwraps2::n_perc0(.data$Pregnancies == 3, show_symbol = TRUE, na_rm = TRUE),
            "4"  = ~ qwraps2::n_perc0(.data$Pregnancies == 4, show_symbol = TRUE, na_rm = TRUE),
            "5"  = ~ qwraps2::n_perc0(.data$Pregnancies == 5, show_symbol = TRUE, na_rm = TRUE),
            "6"  = ~ qwraps2::n_perc0(.data$Pregnancies == 6, show_symbol = TRUE, na_rm = TRUE),
            "7"  = ~ qwraps2::n_perc0(.data$Pregnancies == 7, show_symbol = TRUE, na_rm = TRUE),
            "8"  = ~ qwraps2::n_perc0(.data$Pregnancies == 8, show_symbol = TRUE, na_rm = TRUE),
            "9"  = ~ qwraps2::n_perc0(.data$Pregnancies == 9, show_symbol = TRUE, na_rm = TRUE),
            "10"  = ~ qwraps2::n_perc0(.data$Pregnancies == 10, show_symbol = TRUE, na_rm = TRUE),
            "11"  = ~ qwraps2::n_perc0(.data$Pregnancies == 11, show_symbol = TRUE, na_rm = TRUE),
            "12"  = ~ qwraps2::n_perc0(.data$Pregnancies == 12, show_symbol = TRUE, na_rm = TRUE),
            "13"  = ~ qwraps2::n_perc0(.data$Pregnancies == 13, show_symbol = TRUE, na_rm = TRUE),
            "14"  = ~ qwraps2::n_perc0(.data$Pregnancies == 14, show_symbol = TRUE, na_rm = TRUE),
            "15"  = ~ qwraps2::n_perc0(.data$Pregnancies == 15, show_symbol = TRUE, na_rm = TRUE),
            "16"  = ~ qwraps2::n_perc0(.data$Pregnancies == 16, show_symbol = TRUE, na_rm = TRUE),
            "17"  = ~ qwraps2::n_perc0(.data$Pregnancies == 17, show_symbol = TRUE, na_rm = TRUE))
       )

table1 <- summary_table(Pima_diabetes, our_summary1)
table1

# group number of pregnancies in categories in the table, add descriptives for other variables: mean & SD, median and IQR for age
our_summary2 <-
  list("Diabetes" =
       list("yes" = ~ qwraps2::n_perc0(.data$Diabetes == "yes", show_symbol = TRUE, na_rm = TRUE),
            "no"  = ~ qwraps2::n_perc0(.data$Diabetes == "no", show_symbol = TRUE, na_rm = TRUE)),
       "Number of pregnancies" =
       list("0-1" = ~ qwraps2::n_perc0(.data$Pregnancies < 2, show_symbol = TRUE, na_rm = TRUE),
            "2-3"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 2:3, show_symbol = TRUE, na_rm = TRUE),
            "4-6"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 4:6, show_symbol = TRUE, na_rm = TRUE),
            "7-17"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 7:17, show_symbol = TRUE, na_rm = TRUE)),
       "Age" =
       list("Median (IQR)" = ~ median_iqr(.data$Age)),
       "BMI" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$BMI, denote_sd = "paren", na_rm = TRUE, show_n = "never")),
       "Blood Pressure" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$BloodPressure, denote_sd = "paren", na_rm = TRUE, show_n = "never")),
       "Glucose" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$Glucose, denote_sd = "paren", na_rm = TRUE, show_n = "never"))
       )
table2 <- summary_table(Pima_diabetes, our_summary2)
table2
```

Quicker code to generate the same table:

```{r, results = "asis"}
library(dplyr)
# This needs to reformat the Pregnancies variable first
Pima_diabetes$Pregnancies2 <- as.factor(Pima_diabetes$Pregnancies)

Pima_diabetes %>%
  dplyr::select(.data$Diabetes, .data$Pregnancies2) %>%
  summary_table(.)
```

Generate stratified tables:

```{r, results = "asis"}
our_summary3 <-
  list("Number of pregnancies" =
       list("0-1" = ~ qwraps2::n_perc0(.data$Pregnancies < 2, show_symbol = TRUE, na_rm = TRUE),
            "2-3"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 2:3, show_symbol = TRUE, na_rm = TRUE),
            "4-6"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 4:6, show_symbol = TRUE, na_rm = TRUE),
            "7-17"  = ~ qwraps2::n_perc0(.data$Pregnancies %in% 7:17, show_symbol = TRUE, na_rm = TRUE)),
       "Age" =
       list("Median (IQR)" = ~ median_iqr(.data$Age)),
       "BMI" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$BMI, denote_sd = "paren", na_rm = TRUE, show_n = "never")),
       "Blood Pressure" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$BloodPressure, denote_sd = "paren", na_rm = TRUE, show_n = "never")),
       "Glucose" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(.data$Glucose, denote_sd = "paren", na_rm = TRUE, show_n = "never"))
       )
table3 <- summary_table(dplyr::group_by(Pima_diabetes, Diabetes), our_summary3)
table3

# add column with total and format a bit:
table3_overall <- summary_table(Pima_diabetes, our_summary3)
table3_both <- cbind(table3_overall, table3)

print(table3_both,
      rtitle = "Summary Statistics",
      cnames = c("Overall \\ (N = 768)", "Diabetes \\ (N = 500)", "No Diabetes \\ (N = 268)"))
```

Create stratified table with **table1()** function in **table1** package:

```{r}
# This only seems to work when knitting to html, not to pdf or word
# install.packages("table1")
library(table1)
table1::table1(~ Pregnancies + Age + BMI + BloodPressure + Glucose | Diabetes, data = Pima_diabetes)
```

# Exercise 2: Tables

a) Add descriptive statistics for the **Insulin** variable to the above table generated by the summary_table function:

...

b) Generate a table including descriptive statistics of diabetes prevalence, age, BMI, blood pressure, glucose levels stratified by number of pregnancy quartiles.

...

# Exercise 3: Plots using ggplot2

a) Overview of standard plots 

```{r}
# Load data
load(file = url("https://www.dropbox.com/s/j50z4e17l5aabk8/Pima_diabetes.RData?dl=1"))
Pima_diabetes$Diabetes <- factor(Pima_diabetes$Outcome, labels = c("no", "yes"))

# install.packages("ggplot2")
library(ggplot2)

# Bar plot
ggplot(data = Pima_diabetes, mapping = aes(x = Pregnancies)) + geom_bar()
# same but in short:
ggplot(Pima_diabetes, aes(Pregnancies)) + geom_bar()

# Histogram
ggplot(Pima_diabetes, aes(BMI)) + geom_histogram()

# Histogram and smoothed kernel density estimate
ggplot(Pima_diabetes, aes(BMI, stat(density))) + geom_histogram() + geom_density(col = "red")

# Stratified smoothed kernel density estimate
ggplot(Pima_diabetes, aes(BMI, col = Diabetes)) + geom_density(position = "stack")
ggplot(Pima_diabetes, aes(BMI, fill = Diabetes)) + geom_density(position = "stack")

# Stratified boxplot
ggplot(Pima_diabetes, aes(Diabetes, BMI)) + geom_boxplot()

# Scatterplot
ggplot(Pima_diabetes, aes(BMI, Glucose)) + geom_point()

# Scatterplot using color for third variable
ggplot(Pima_diabetes, aes(BMI, Glucose, col = Diabetes)) + geom_point()

# A bit more fancy and complicated scatterplots:
ggplot(Pima_diabetes, aes(BMI, Glucose, col = Pregnancies)) + 
  geom_point() + 
  scale_colour_gradientn(colours = rainbow(4)) +
  facet_wrap(~Diabetes)

Pima_diabetes$ManyPregnancies <- Pima_diabetes$Pregnancies>=4
Pima_diabetes$ManyPregnancies <- factor(Pima_diabetes$ManyPregnancies, labels = c("< 4 Pregnancies", ">= 4 Pregnancies"))
ggplot(Pima_diabetes, aes(BMI, Glucose, col = ManyPregnancies)) + 
  geom_point() +
  facet_wrap(~Diabetes)
```

b) Generate the following plots

* Produce a boxplot of BMI for each group of number of pregnancies
* In the above scatterplot of BMI and glucose levels, add a regression line which describes the linear association between the two variables.

...

# Exercise 4: Other cool packages and graphics in R

The following two examples can only be knitted to html.

Interactive maps: 

```{r}
# install.packages("leaflet")
library(leaflet)

m <- leaflet() %>% addTiles()

m %>%
  # Central Park
  fitBounds(-73.9, 40.75, -73.95,40.8) %>%
  addMeasure()

m %>%
  # Berlin, Germany
  fitBounds(13.76134, 52.675499, 13.0884, 52.33812) %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479")
```

Interactive ggplot2 plots:

```{r}
# install.packages("plotly")
library(plotly)

plot_BMI_glu_diab <- ggplot(Pima_diabetes, aes(BMI, Glucose, col = Diabetes)) + geom_point()
plot_BMI_glu_diab_interactive <- ggplotly(plot_BMI_glu_diab)

plot_BMI_glu_diab_interactive
```
