---
title: "Homework 4"
author: "Stefan Konigorski"
date: "November 11, 2019"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Descriptive statistics - example 1

Create a summary table with descriptive statistics of the esoph dataframe in the datasets package. For example by using the summary_table function in the qwraps package. Alternatively, only compute the descriptive statistics of interest.

```{r}
library(datasets)
?esoph
dat1 <- esoph
```

## Exercise 2: Descriptive statistics - example 2

Create a summary table with descriptive statistics of the UKLungDeaths data in the datasets package. For example by using the summary_table function in the qwraps package. Alternatively, only compute the descriptive statistics of interest.

```{r}
library(datasets)
?UKLungDeaths 
```

Hint: the data consists of three datasets.
