---
title: "Missing values - Introductory example"
author: "Stefan Konigorski"
date: "December 2, 2019"
output:
  html_document: default
---

## Load KiGGS data

```{r}
dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
load(dat_link)

# Alternative: download data from Moodle, either (i) the raw .sav dataset (SPSS data file) which you need to import, or (ii) the already imported .RData file with:
# load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
# (you need to adapt path)

dat <- KiGGS03_06
```

## Example of possible situation with missing values

Study question: 

Is there an association between the length of nightly sleep of children and their psychological wellbeing?

Variables available to investigate this study question:

```{r}
# sleep:
str(dat$e065z)
head(dat$e065z)
summary(dat$e065z)

# Psychological wellbeing of the children (self rating)
str(dat$pw100.k)
head(dat$pw100.k)
summary(dat$pw100.k)
```

Number of missing values:

```{r}
table(!is.na(as.numeric(as.character(dat$pw100.k))))
table(!is.na(as.numeric(as.character(dat$e065z))))

table(!is.na(as.numeric(as.character(dat$pw100.k))) & !is.na(as.numeric(as.character(dat$e065z))))
```

Question: What is the best way to investigate the study question?
-> e.g. use the rating of the children's wellbeing by their parents

```{r}
# Psychological wellbeing of the children (external rating by parents)
summary(dat$pw100.e)
table(!is.na(as.numeric(as.character(dat$pw100.e))))

table(!is.na(as.numeric(as.character(dat$pw100.e))) & !is.na(as.numeric(as.character(dat$e065z))))
```

