---
title: "How to deal with missing values"
author: "Stefan Konigorski"
date: "December 2, 2019"
output:
  html_document: default
---

## Overview

Here, in the analysis, the aim is to investigate the association between the length of nightly sleep of children and their psychological wellbeing. 

Both variables are quantitative variables. For the psychological wellbeing of children, the external rating by their parents is used (see R_7a_missing_values_intro_example.Rmd). The sleep variable is approx. normally distributed, the wellbeing variable is not normally distributed. Still, we will use the Pearson correlation coefficient for the analysis here (which is most well-suited for the analysis normally distributed variables). Therefore, the results have to be interpreted with some caution.

In a first step, we will create datasets that have missing values according to different underlying mechanisms:

1. Full dataset with all observations of sleep and wellbeing.
2. Values of psychological wellbeing are deleted randomly (10\%, 50\%, 90\%) -> MCAR
3. Values of psychological wellbeing are deleted (10\%, 50\%, 90\%) depending on values of physical wellbeing. -> MAR
4. Values of psychological wellbeing are deleted (10\%, 50\%, 90\%) depending on values of psychological wellbeing itself. -> MNAR

In the next step, we will analyze these different datasets with different methods dealing with the missing values, and we will investigate how biased the estimated correlations are, and if the results (power) of the hypothesis test changes.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/nbh0f6upejh7nr8/data.RData?dl=1")
load(dat_link)

# Alternative: download data from Moodle, either (i) the raw .sav dataset (SPSS data file) which you need to import, or (ii) the already imported .RData file with:
# load("C:/Users/stefan.konigorski/Desktop/data_Teaching/KiGGS/KiGGS03_06.RData")
# (you need to adapt the path)

dat <- KiGGS03_06
```

Overview of the variables:

```{r}
# sleep:
str(dat$e065z)
summary(dat$e065z)
hist(as.numeric(as.character(dat$e065z)))

# Psychological wellbeing of the children (external rating by parents)
str(dat$pw100.e)
summary(dat$pw100.e)
hist(as.numeric(as.character(dat$pw100.e)))

# Physical wellbeing of the children (external rating by parents)
str(dat$kw100.e)
summary(dat$kw100.e)
hist(as.numeric(as.character(dat$kw100.e)))
```

## Create datasets

```{r}
# Full dataset:
dat_complete <- data.frame(schlaf = as.numeric(as.character(dat$e065z)), 
                           psy = as.numeric(as.character(dat$pw100.e)), 
                           body = as.numeric(as.character(dat$kw100.e)))
idx_complete <- complete.cases(dat_complete)
dat_complete <- dat_complete[idx_complete, ]

table(!is.na(dat_complete$psy))

# Datasets in which 10%, 50%, 90% of the observations of psychological wellbeing are deleted randomly
n_total <- dim(dat_complete)[1]

set.seed(5)
idx_MCAR_10 <- sample(1:n_total, size = 0.1*n_total, replace = FALSE)
set.seed(5)
idx_MCAR_50 <- sample(1:n_total, size = 0.5*n_total, replace = FALSE)
set.seed(5)
idx_MCAR_90 <- sample(1:n_total, size = 0.9*n_total, replace = FALSE)

dat_MCAR_10 <- dat_MCAR_50 <- dat_MCAR_90 <- dat_complete
dat_MCAR_10$psy[idx_MCAR_10] <- NA
dat_MCAR_50$psy[idx_MCAR_50] <- NA
dat_MCAR_90$psy[idx_MCAR_90] <- NA

table(!is.na(dat_MCAR_10$psy))
table(!is.na(dat_MCAR_50$psy))
table(!is.na(dat_MCAR_90$psy))

# Datasets in which 10%, 50%, 90% of the observations of psychological wellbeing are deleted depending on values of physical wellbeing.

quantile(dat_complete$body, probs = c(0.1, 0.5, 0.9), na.rm = TRUE)

idx_MAR_10 <- (dat_complete$body >= 100)
idx_MAR_50 <- (dat_complete$body > 81.25)
idx_MAR_90 <- (dat_complete$body >= 62.5)

dat_MAR_10 <- dat_MAR_50 <- dat_MAR_90 <- dat_complete
dat_MAR_10$psy[idx_MAR_10] <- NA
dat_MAR_50$psy[idx_MAR_50] <- NA
dat_MAR_90$psy[idx_MAR_90] <- NA

table(!is.na(dat_MAR_10$psy))
table(!is.na(dat_MAR_50$psy))
table(!is.na(dat_MAR_90$psy))

# Information on the association of physical and psychological wellbeing and sleep:
cor(dat_complete$body, dat_complete$psy)
cor(dat_complete$body, dat_complete$schlaf)

# Datasets in which 10%, 50%, 90% of the observations of psychological wellbeing are deleted depending on the values of psychological wellbeing itself.

quantile(dat_complete$psy, probs = c(0.1, 0.5, 0.9), na.rm = TRUE)

idx_MNAR_10 <- (dat_complete$psy < 68.75)
idx_MNAR_50 <- (dat_complete$psy < 87.50)
idx_MNAR_90 <- (dat_complete$psy <= 93.75)

dat_MNAR_10 <- dat_MNAR_50 <- dat_MNAR_90 <- dat_complete
dat_MNAR_10$psy[idx_MNAR_10] <- NA
dat_MNAR_50$psy[idx_MNAR_50] <- NA
dat_MNAR_90$psy[idx_MNAR_90] <- NA

table(!is.na(dat_MNAR_10$psy))
table(!is.na(dat_MNAR_50$psy))
table(!is.na(dat_MNAR_90$psy))
```

## Ground Truth: true correlation (=measure of association), 95% confidence interval, and p-value of hypothesis test

```{r}
res0 <- cor.test(dat_complete$schlaf, dat_complete$psy, method = "pearson")
res0$estimate
res0$conf.int
res0$p.value
```

## Analysis of the different datasets with missing values

### Complete-Case analysis

```{r}
# MCAR:
res1_10 <- cor.test(dat_MCAR_10$schlaf, dat_MCAR_10$psy, method = "pearson")
res1_10$estimate
res1_10$conf.int
res1_10$p.value

res1_50 <- cor.test(dat_MCAR_50$schlaf, dat_MCAR_50$psy, method = "pearson")
res1_50$estimate
res1_50$conf.int
res1_50$p.value

res1_90 <- cor.test(dat_MCAR_90$schlaf, dat_MCAR_90$psy, method = "pearson")
res1_90$estimate
res1_90$conf.int
res1_90$p.value
```

Conclusion: Unbiased estimates, confidence intervals are wider, the more missing values there are (sample size of analysis is getting smaller).

```{r}
# MAR:
res2_10 <- cor.test(dat_MAR_10$schlaf, dat_MAR_10$psy, method = "pearson")
res2_10$estimate
res2_10$conf.int
res2_10$p.value

res2_50 <- cor.test(dat_MAR_50$schlaf, dat_MAR_50$psy, method = "pearson")
res2_50$estimate
res2_50$conf.int
res2_50$p.value

res2_90 <- cor.test(dat_MAR_90$schlaf, dat_MAR_90$psy, method = "pearson")
res2_90$estimate
res2_90$conf.int
res2_90$p.value
```

Conclusion: Biased estimates - the more missing values, the more biased. Also, the confidence intervals are getting wider with increasing number of missing values.

```{r}
# MNAR:
res3_10 <- cor.test(dat_MNAR_10$schlaf, dat_MNAR_10$psy, method = "pearson")
res3_10$estimate
res3_10$conf.int
res3_10$p.value

res3_50 <- cor.test(dat_MNAR_50$schlaf, dat_MNAR_50$psy, method = "pearson")
res3_50$estimate
res3_50$conf.int
res3_50$p.value

res3_90 <- cor.test(dat_MNAR_90$schlaf, dat_MNAR_90$psy, method = "pearson")
res3_90$estimate
res3_90$conf.int
res3_90$p.value
```

Conclusion: Biased estimates - the more missing values, the more biased. Also, the confidence intervals are getting wider with increasing number of missing values.

### Mean imputation

```{r}
# MCAR:
mean(dat_MCAR_10$psy, na.rm = TRUE)
dat_MCAR_10_mean <- dat_MCAR_10
dat_MCAR_10_mean$psy[is.na(dat_MCAR_10_mean$psy)] <- mean(dat_MCAR_10_mean$psy, na.rm = TRUE)
res1_10 <- cor.test(dat_MCAR_10_mean$schlaf, dat_MCAR_10_mean$psy, method = "pearson")
res1_10$estimate
res1_10$conf.int
res1_10$p.value

mean(dat_MCAR_50$psy, na.rm = TRUE)
dat_MCAR_50_mean <- dat_MCAR_50
dat_MCAR_50_mean$psy[is.na(dat_MCAR_50_mean$psy)] <- mean(dat_MCAR_50_mean$psy, na.rm = TRUE)
res1_50 <- cor.test(dat_MCAR_50_mean$schlaf, dat_MCAR_50_mean$psy, method = "pearson")
res1_50$estimate
res1_50$conf.int
res1_50$p.value

mean(dat_MCAR_90$psy, na.rm = TRUE)
dat_MCAR_90_mean <- dat_MCAR_90
dat_MCAR_90_mean$psy[is.na(dat_MCAR_90_mean$psy)] <- mean(dat_MCAR_90_mean$psy, na.rm = TRUE)
res1_90 <- cor.test(dat_MCAR_90_mean$schlaf, dat_MCAR_90_mean$psy, method = "pearson")
res1_90$estimate
res1_90$conf.int
res1_90$p.value
```

Conclusion: ...

```{r}
# MAR:
mean(dat_MAR_10$psy, na.rm = TRUE)
dat_MAR_10_mean <- dat_MAR_10
dat_MAR_10_mean$psy[is.na(dat_MAR_10_mean$psy)] <- mean(dat_MAR_10_mean$psy, na.rm = TRUE)
res2_10 <- cor.test(dat_MAR_10_mean$schlaf, dat_MAR_10_mean$psy, method = "pearson")
res2_10$estimate
res2_10$conf.int
res2_10$p.value

mean(dat_MAR_50$psy, na.rm = TRUE)
dat_MAR_50_mean <- dat_MAR_50
dat_MAR_50_mean$psy[is.na(dat_MAR_50_mean$psy)] <- mean(dat_MAR_50_mean$psy, na.rm = TRUE)
res2_50 <- cor.test(dat_MAR_50_mean$schlaf, dat_MAR_50_mean$psy, method = "pearson")
res2_50$estimate
res2_50$conf.int
res2_50$p.value

mean(dat_MAR_90$psy, na.rm = TRUE)
dat_MAR_90_mean <- dat_MAR_90
dat_MAR_90_mean$psy[is.na(dat_MAR_90_mean$psy)] <- mean(dat_MAR_90_mean$psy, na.rm = TRUE)
res2_90 <- cor.test(dat_MAR_90_mean$schlaf, dat_MAR_90_mean$psy, method = "pearson")
res2_90$estimate
res2_90$conf.int
res2_90$p.value
```

Conclusion: ...

```{r}
# MNAR:
mean(dat_MNAR_10$psy, na.rm = TRUE)
dat_MNAR_10_mean <- dat_MNAR_10
dat_MNAR_10_mean$psy[is.na(dat_MNAR_10_mean$psy)] <- mean(dat_MNAR_10_mean$psy, na.rm = TRUE)
res3_10 <- cor.test(dat_MNAR_10_mean$schlaf, dat_MNAR_10_mean$psy, method = "pearson")
res3_10$estimate
res3_10$conf.int
res3_10$p.value

mean(dat_MNAR_50$psy, na.rm = TRUE)
dat_MNAR_50_mean <- dat_MNAR_50
dat_MNAR_50_mean$psy[is.na(dat_MNAR_50_mean$psy)] <- mean(dat_MNAR_50_mean$psy, na.rm = TRUE)
res3_50 <- cor.test(dat_MNAR_50_mean$schlaf, dat_MNAR_50_mean$psy, method = "pearson")
res3_50$estimate
res3_50$conf.int
res3_50$p.value

mean(dat_MNAR_90$psy, na.rm = TRUE)
dat_MNAR_90_mean <- dat_MNAR_90
dat_MNAR_90_mean$psy[is.na(dat_MNAR_90_mean$psy)] <- mean(dat_MNAR_90_mean$psy, na.rm = TRUE)
res3_90 <- cor.test(dat_MNAR_90_mean$schlaf, dat_MNAR_90_mean$psy, method = "pearson")
res3_90$estimate
res3_90$conf.int
res3_90$p.value
```

Conclusion: ...

### Median imputation

```{r}
# MCAR:
median(dat_MCAR_10$psy, na.rm = TRUE)
dat_MCAR_10_median <- dat_MCAR_10
dat_MCAR_10_median$psy[is.na(dat_MCAR_10_median$psy)] <- median(dat_MCAR_10_median$psy, na.rm = TRUE)
res1_10 <- cor.test(dat_MCAR_10_median$schlaf, dat_MCAR_10_median$psy, method = "pearson")
res1_10$estimate
res1_10$conf.int
res1_10$p.value

median(dat_MCAR_50$psy, na.rm = TRUE)
dat_MCAR_50_median <- dat_MCAR_50
dat_MCAR_50_median$psy[is.na(dat_MCAR_50_median$psy)] <- median(dat_MCAR_50_median$psy, na.rm = TRUE)
res1_50 <- cor.test(dat_MCAR_50_median$schlaf, dat_MCAR_50_median$psy, method = "pearson")
res1_50$estimate
res1_50$conf.int
res1_50$p.value

median(dat_MCAR_90$psy, na.rm = TRUE)
dat_MCAR_90_median <- dat_MCAR_90
dat_MCAR_90_median$psy[is.na(dat_MCAR_90_median$psy)] <- median(dat_MCAR_90_median$psy, na.rm = TRUE)
res1_90 <- cor.test(dat_MCAR_90_median$schlaf, dat_MCAR_90_median$psy, method = "pearson")
res1_90$estimate
res1_90$conf.int
res1_90$p.value
```

Conclusion: ...

```{r}
# MAR:
median(dat_MAR_10$psy, na.rm = TRUE)
dat_MAR_10_median <- dat_MAR_10
dat_MAR_10_median$psy[is.na(dat_MAR_10_median$psy)] <- median(dat_MAR_10_median$psy, na.rm = TRUE)
res2_10 <- cor.test(dat_MAR_10_median$schlaf, dat_MAR_10_median$psy, method = "pearson")
res2_10$estimate
res2_10$conf.int
res2_10$p.value

median(dat_MAR_50$psy, na.rm = TRUE)
dat_MAR_50_median <- dat_MAR_50
dat_MAR_50_median$psy[is.na(dat_MAR_50_median$psy)] <- median(dat_MAR_50_median$psy, na.rm = TRUE)
res2_50 <- cor.test(dat_MAR_50_median$schlaf, dat_MAR_50_median$psy, method = "pearson")
res2_50$estimate
res2_50$conf.int
res2_50$p.value

median(dat_MAR_90$psy, na.rm = TRUE)
dat_MAR_90_median <- dat_MAR_90
dat_MAR_90_median$psy[is.na(dat_MAR_90_median$psy)] <- median(dat_MAR_90_median$psy, na.rm = TRUE)
res2_90 <- cor.test(dat_MAR_90_median$schlaf, dat_MAR_90_median$psy, method = "pearson")
res2_90$estimate
res2_90$conf.int
res2_90$p.value
```

Conclusion: ...

```{r}
# MNAR:
median(dat_MNAR_10$psy, na.rm = TRUE)
dat_MNAR_10_median <- dat_MNAR_10
dat_MNAR_10_median$psy[is.na(dat_MNAR_10_median$psy)] <- median(dat_MNAR_10_median$psy, na.rm = TRUE)
res3_10 <- cor.test(dat_MNAR_10_median$schlaf, dat_MNAR_10_median$psy, method = "pearson")
res3_10$estimate
res3_10$conf.int
res3_10$p.value

median(dat_MNAR_50$psy, na.rm = TRUE)
dat_MNAR_50_median <- dat_MNAR_50
dat_MNAR_50_median$psy[is.na(dat_MNAR_50_median$psy)] <- median(dat_MNAR_50_median$psy, na.rm = TRUE)
res3_50 <- cor.test(dat_MNAR_50_median$schlaf, dat_MNAR_50_median$psy, method = "pearson")
res3_50$estimate
res3_50$conf.int
res3_50$p.value

median(dat_MNAR_90$psy, na.rm = TRUE)
dat_MNAR_90_median <- dat_MNAR_90
dat_MNAR_90_median$psy[is.na(dat_MNAR_90_median$psy)] <- median(dat_MNAR_90_median$psy, na.rm = TRUE)
res3_90 <- cor.test(dat_MNAR_90_median$schlaf, dat_MNAR_90_median$psy, method = "pearson")
res3_90$estimate
res3_90$conf.int
res3_90$p.value
```

Conclusion: ...

### Regression-based imputation

```{r}
# MCAR:
reg_1_10 <- lm(psy ~ body, data = dat_MCAR_10)
dat_MCAR_10_reg <- dat_MCAR_10
dat_MCAR_10_reg$psy[is.na(dat_MCAR_10_reg$psy)] <- coef(reg_1_10)[1] + coef(reg_1_10)[2] * dat_MCAR_10_reg$body[is.na(dat_MCAR_10_reg$psy)]
res1_10 <- cor.test(dat_MCAR_10_reg$schlaf, dat_MCAR_10_reg$psy, method = "pearson")
res1_10$estimate
res1_10$conf.int
res1_10$p.value

reg_1_50 <- lm(psy ~ body, data = dat_MCAR_50)
dat_MCAR_50_reg <- dat_MCAR_50
dat_MCAR_50_reg$psy[is.na(dat_MCAR_50_reg$psy)] <- coef(reg_1_50)[1] + coef(reg_1_50)[2] * dat_MCAR_50_reg$body[is.na(dat_MCAR_50_reg$psy)]
res1_50 <- cor.test(dat_MCAR_50_reg$schlaf, dat_MCAR_50_reg$psy, method = "pearson")
res1_50$estimate
res1_50$conf.int
res1_50$p.value

reg_1_90 <- lm(psy ~ body, data = dat_MCAR_90)
dat_MCAR_90_reg <- dat_MCAR_90
dat_MCAR_90_reg$psy[is.na(dat_MCAR_90_reg$psy)] <- coef(reg_1_90)[1] + coef(reg_1_90)[2] * dat_MCAR_90_reg$body[is.na(dat_MCAR_90_reg$psy)]
res1_90 <- cor.test(dat_MCAR_90_reg$schlaf, dat_MCAR_90_reg$psy, method = "pearson")
res1_90$estimate
res1_90$conf.int
res1_90$p.value
```

Conclusion: ...

```{r}
# MAR:
reg_2_10 <- lm(psy ~ body, data = dat_MAR_10)
dat_MAR_10_reg <- dat_MAR_10
dat_MAR_10_reg$psy[is.na(dat_MAR_10_reg$psy)] <- coef(reg_2_10)[1] + coef(reg_2_10)[2] * dat_MAR_10_reg$body[is.na(dat_MAR_10_reg$psy)]
res2_10 <- cor.test(dat_MAR_10_reg$schlaf, dat_MAR_10_reg$psy, method = "pearson")
res2_10$estimate
res2_10$conf.int
res2_10$p.value

reg_2_50 <- lm(psy ~ body, data = dat_MAR_50)
dat_MAR_50_reg <- dat_MAR_50
dat_MAR_50_reg$psy[is.na(dat_MAR_50_reg$psy)] <- coef(reg_2_50)[1] + coef(reg_2_50)[2] * dat_MAR_50_reg$body[is.na(dat_MAR_50_reg$psy)]
res2_50 <- cor.test(dat_MAR_50_reg$schlaf, dat_MAR_50_reg$psy, method = "pearson")
res2_50$estimate
res2_50$conf.int
res2_50$p.value

reg_2_90 <- lm(psy ~ body, data = dat_MAR_90)
dat_MAR_90_reg <- dat_MAR_90
dat_MAR_90_reg$psy[is.na(dat_MAR_90_reg$psy)] <- coef(reg_2_90)[1] + coef(reg_2_90)[2] * dat_MAR_90_reg$body[is.na(dat_MAR_90_reg$psy)]
res2_90 <- cor.test(dat_MAR_90_reg$schlaf, dat_MAR_90_reg$psy, method = "pearson")
res2_90$estimate
res2_90$conf.int
res2_90$p.value
```

Conclusion: ...

```{r}
# MNAR:
reg_3_10 <- lm(psy ~ body, data = dat_MNAR_10)
dat_MNAR_10_reg <- dat_MNAR_10
dat_MNAR_10_reg$psy[is.na(dat_MNAR_10_reg$psy)] <- coef(reg_3_10)[1] + coef(reg_3_10)[2] * dat_MNAR_10_reg$body[is.na(dat_MNAR_10_reg$psy)]
res3_10 <- cor.test(dat_MNAR_10_reg$schlaf, dat_MNAR_10_reg$psy, method = "pearson")
res3_10$estimate
res3_10$conf.int
res3_10$p.value

reg_3_50 <- lm(psy ~ body, data = dat_MNAR_50)
dat_MNAR_50_reg <- dat_MNAR_50
dat_MNAR_50_reg$psy[is.na(dat_MNAR_50_reg$psy)] <- coef(reg_3_50)[1] + coef(reg_3_50)[2] * dat_MNAR_50_reg$body[is.na(dat_MNAR_50_reg$psy)]
res3_50 <- cor.test(dat_MNAR_50_reg$schlaf, dat_MNAR_50_reg$psy, method = "pearson")
res3_50$estimate
res3_50$conf.int
res3_50$p.value

reg_3_90 <- lm(psy ~ body, data = dat_MNAR_90)
dat_MNAR_90_reg <- dat_MNAR_90
dat_MNAR_90_reg$psy[is.na(dat_MNAR_90_reg$psy)] <- coef(reg_3_90)[1] + coef(reg_3_90)[2] * dat_MNAR_90_reg$body[is.na(dat_MNAR_90_reg$psy)]
res3_90 <- cor.test(dat_MNAR_90_reg$schlaf, dat_MNAR_90_reg$psy, method = "pearson")
res3_90$estimate
res3_90$conf.int
res3_90$p.value
```

Conclusion: ...

### Multiple imputation using mice

```{r}
# install.packages("mice")
library(mice)

# Pattern of missing values
md.pattern(dat_MCAR_10)

# Nice visualization of missing values in the VIM package, can be useful to identify patterns of missing values
# install.packages("VIM")
library(VIM)
aggr(dat_MCAR_10, col = c('navyblue','red'), numbers = TRUE, sortVars = TRUE, labels = names(dat_MCAR_10), cex.axis = 0.7, gap = 3, ylab = c("Histogram of missing data", "Pattern"))

# Other example how to apply the function
aggr(dat[, 100:110])
```

```{r, include = FALSE}
# Step 1: Do imputation of missing values, generate 5 imputed datasets
tempData <- mice(dat_MCAR_10, m = 5, maxit = 5, meth = 'pmm', seed = 500)
?mice
```

```{r}
# Summary of the imputation
summary(tempData)

# The 5 imputed datasets can be extracted as follows:
completedData1 <- complete(tempData,1)
completedData2 <- complete(tempData,2)
completedData3 <- complete(tempData,3)
completedData4 <- complete(tempData,4)
completedData5 <- complete(tempData,5)

# Visualization of the distribution of the imputed variable psy
mice::densityplot(tempData)

# Step 2: Do the statistical analysis using the "with" function
modelFit1 <- with(tempData, cor(psy, schlaf, method = "pearson"))

# The estimate of the correlation can also be calculated by hand:
mean(unlist(modelFit1$analyses))

# Alternatively, you can use the cor.test function, which also computes confidence intervals
with(tempData, cor.test(psy, schlaf, method = "pearson"))

# Step 3: Pool results
# This can be done using the "pool" function in the mice package, where dfcom are the degrees of freedom of the separate analyses of the imputed datasets, which are also shown in the cor.test output. eg:
# summary(pool(modelFit1), dfcom = 7557)
# However, the pooling of the correlations i.e. the extraction of the relevant information is not readily available in the pool function, so that this gives an error here.

# For regression it works, e.g. using:
modelFit1 <- with(tempData, lm(psy ~ schlaf))
summary(pool(modelFit1), dfcom = 7557)

# For correlations you can do this by hand, e.g. by extracting the estimators and their standard errors from the results, for the pooling,
# e.g. do something like modelFit1_for_pool <- list(est = ..., SE = ...)

# Simpler: use the miceadds package:
# install.packages("miceadds")
library(miceadds)
```

```{r, include = FALSE}
tempData <- mice(dat_MCAR_10, m = 5, maxit = 5, meth = 'pmm', seed = 500)
```

```{r}
res <- miceadds::micombine.cor(mi.res = tempData, variables = c(1,2), conf.level = 0.95, method = "pearson")
```

```{r, include = FALSE}
# In summary:
library(mice)
library(miceadds)

tempData_1_10 <- mice(dat_MCAR_10, meth = 'pmm', seed = 500)
tempData_1_50 <- mice(dat_MCAR_50, meth = 'pmm', seed = 500)
tempData_1_90 <- mice(dat_MCAR_90, meth = 'pmm', seed = 500)
tempData_2_10 <- mice(dat_MAR_10, meth = 'pmm', seed = 500)
tempData_2_50 <- mice(dat_MAR_50, meth = 'pmm', seed = 500)
tempData_2_90 <- mice(dat_MAR_90, meth = 'pmm', seed = 500)
tempData_3_10 <- mice(dat_MNAR_10, meth = 'pmm', seed = 500)
tempData_3_50 <- mice(dat_MNAR_50, meth = 'pmm', seed = 500)
tempData_3_90 <- mice(dat_MNAR_90, meth = 'pmm', seed = 500)
```

```{r}
# MCAR:
res1_10 <- miceadds::micombine.cor(mi.res = tempData_1_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_10$r[1]
c(res1_10$lower95[1], res1_10$upper95[1]) #pooled 95% CI
res1_10$p[1]

res1_50 <- miceadds::micombine.cor(mi.res = tempData_1_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_50$r[1]
c(res1_50$lower95[1], res1_50$upper95[1]) #pooled 95% CI
res1_50$p[1]

res1_90 <- miceadds::micombine.cor(mi.res = tempData_1_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_90$r[1]
c(res1_90$lower95[1], res1_90$upper95[1]) #pooled 95% CI
res1_90$p[1]

# MAR:
res2_10 <- miceadds::micombine.cor(mi.res = tempData_2_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_10$r[1]
c(res2_10$lower95[1], res2_10$upper95[1]) #pooled 95% CI
res2_10$p[1]

res2_50 <- miceadds::micombine.cor(mi.res = tempData_2_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_50$r[1]
c(res2_50$lower95[1], res2_50$upper95[1]) #pooled 95% CI
res2_50$p[1]

res2_90 <- miceadds::micombine.cor(mi.res = tempData_2_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_90$r[1]
c(res2_90$lower95[1], res2_90$upper95[1]) #pooled 95% CI
res2_90$p[1]

# MNAR:
res3_10 <- miceadds::micombine.cor(mi.res = tempData_3_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_10$r[1]
c(res3_10$lower95[1], res3_10$upper95[1]) #pooled 95% CI
res3_10$p[1]

res3_50 <- miceadds::micombine.cor(mi.res = tempData_3_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_50$r[1]
c(res3_50$lower95[1], res3_50$upper95[1]) #pooled 95% CI
res3_50$p[1]

res3_90 <- miceadds::micombine.cor(mi.res = tempData_3_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_90$r[1]
c(res3_90$lower95[1], res3_90$upper95[1]) #pooled 95% CI
res3_90$p[1]
```

```{r, include = FALSE}
# Using Bayesian linear regression in the imputation:
tempData_1_10 <- mice(dat_MCAR_10, meth = 'norm', seed = 500)
tempData_1_50 <- mice(dat_MCAR_50, meth = 'norm', seed = 500)
tempData_1_90 <- mice(dat_MCAR_90, meth = 'norm', seed = 500)
tempData_2_10 <- mice(dat_MAR_10, meth = 'norm', seed = 500)
tempData_2_50 <- mice(dat_MAR_50, meth = 'norm', seed = 500)
tempData_2_90 <- mice(dat_MAR_90, meth = 'norm', seed = 500)
tempData_3_10 <- mice(dat_MNAR_10, meth = 'norm', seed = 500)
tempData_3_50 <- mice(dat_MNAR_50, meth = 'norm', seed = 500)
tempData_3_90 <- mice(dat_MNAR_90, meth = 'norm', seed = 500)
```

```{r}
# MCAR:
res1_10 <- miceadds::micombine.cor(mi.res = tempData_1_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_10$r[1]
c(res1_10$lower95[1], res1_10$upper95[1]) #pooled 95% CI
res1_10$p[1]

res1_50 <- miceadds::micombine.cor(mi.res = tempData_1_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_50$r[1]
c(res1_50$lower95[1], res1_50$upper95[1]) #pooled 95% CI
res1_50$p[1]

res1_90 <- miceadds::micombine.cor(mi.res = tempData_1_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res1_90$r[1]
c(res1_90$lower95[1], res1_90$upper95[1]) #pooled 95% CI
res1_90$p[1]

# MAR:
res2_10 <- miceadds::micombine.cor(mi.res = tempData_2_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_10$r[1]
c(res2_10$lower95[1], res2_10$upper95[1]) #pooled 95% CI
res2_10$p[1]

res2_50 <- miceadds::micombine.cor(mi.res = tempData_2_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_50$r[1]
c(res2_50$lower95[1], res2_50$upper95[1]) #pooled 95% CI
res2_50$p[1]

res2_90 <- miceadds::micombine.cor(mi.res = tempData_2_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res2_90$r[1]
c(res2_90$lower95[1], res2_90$upper95[1]) #pooled 95% CI
res2_90$p[1]

# MNAR:
res3_10 <- miceadds::micombine.cor(mi.res = tempData_3_10, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_10$r[1]
c(res3_10$lower95[1], res3_10$upper95[1]) #pooled 95% CI
res3_10$p[1]

res3_50 <- miceadds::micombine.cor(mi.res = tempData_3_50, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_50$r[1]
c(res3_50$lower95[1], res3_50$upper95[1]) #pooled 95% CI
res3_50$p[1]

res3_90 <- miceadds::micombine.cor(mi.res = tempData_3_90, variables = c(1,2), conf.level = 0.95, method = "pearson")
res3_90$r[1]
c(res3_90$lower95[1], res3_90$upper95[1]) #pooled 95% CI
res3_90$p[1]
```
